---
title: "Dissertation Chapter 3 Statistical Analysis Models"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
```

## Chapter 3 AD RELATED PATHOLOGY RELATES TO FUNCTIONAL BRAIN NETWORK ORGANIZATION

```{r define covariates}
# define covariates for cross-sectional analyses:
# age, self-reported gender,education status, head motion (post-scrubbing FD)
cross_cov = 'Age+Gender+Educ_Group+FD_postscrub'

# define covariates for longitudinal analyses: 
# age at baseline, self-reported gender,education status, head motion (post-scrubbing FD) at baseline
long_cov = 'Age_bl+Gender+Educ_Group+FD_postscrub_bl'
```

All reported cross-sectional analyses controlled for the above cross-sectional analyses covariates. 
All reported longitudinal analyses controlled for the above longitudinal analyses covariates.

### 3.3.2. Aβ burden is not related to brain system segregation

1) Examine the relationship between Centiloids and brain system segregation: 

```{r model 3.3.2.1}
# DV: whole-brain system segregation
# IV: Centiloids
mod3.3.2.1 <- paste('wb_seg ~ Centiloids',cross_cov,sep = "+") 

 # perform multiple linear regression: 
lm(formula = mod3.3.2.1,data = data, 
   contrasts = list(Gender="contr.sum",Educ_Group=c(-1,1))) 
```

2)  Examine whether there exists a difference in brain system segregation across cortical Aβ groups (Aβ+ vs. Aβ-): 

```{r model 3.3.2.2}
# DV: whole-brain system segregation
# IV: Aβ burden status (Aβ+ vs. Aβ-)

mod3.3.2.2 <- paste('wb_seg ~ Amyloid_Group',cross_cov,sep = "+") 

# Perform ANCOVA
car::Anova(lm(formula = mod3.3.2.2,data = data,contrasts=
              list(Amyloid_Group="contr.treatment",Gender = "contr.sum",Educ_Factor=c(-1,1))),type=3) # perform ANCOVA
```

### 3.3.3. Tau burden relates to brain system segregation

1) Examine the relationship between meta-temporal tau and brain system segregation: 

```{r model 3.3.3.1}
# DV: whole-brain system segregation
# IV: meta-temporal tau SUVR

mod3.3.3.1 = paste('wb_seg ~ meta_temporal_suvr',cross_cov,sep = "+")

#perform multiple linear regression
lm(formula = mod3.3.3.1,data = data,contrasts = list(Gender = "contr.sum",Educ_Group=c(-1,1)))
```

2)  Examine whether there exists a difference in brain system segregation across tau groups (Tau+ vs. Tau-): 

```{r model 3.3.3.2}
# DV: whole-brain system segregation
# IV: meta-temporal tau burden status (tau+ vs. tau-)

mod3.3.3.2 = paste('wb_seg ~ Tau_Group',cross_cov,sep = "+") 

# Perform ANCOVA
car::Anova(lm(formula = mod3.3.3.2,data = data,contrasts=
              list(Tau_Group="contr.treatment",Gender = "contr.sum",Educ_Factor=c(-1,1))),type=3)
```

3) Examine the combined effects of Aβ and tau burden on brain system segregation: 

- Examine the interaction between Centiloids and meta-temporal tau uptake on whole-brain system segregation:

```{r model 3.3.3.3}
# DV: whole-brain system segregation
# IV: the interaction between centiloids and meta-temporal tau burden

mod3.3.3.3 = paste('wb_seg ~ Centiloids*meta_temporal_suvr',cross_cov,sep = "+")

# perform multiple linear regression:
lm(formula = mod3.3.3.3,data=data,contrasts = list(Gender = "contr.sum",Educ_Group=c(-1,1)))
```

- Examine the interaction between Centiloids, meta-temporal tau and age on whole-brain system segregation:

```{r model 3.3.3.4}
# DV: whole-brain system segregation
# IV: the interaction between centiloids, meta-temporal tau burden and age
# the covariates for this model specifically is gender,education status, and head motion

mod3.3.3.4 = paste('wb_seg ~ centiloids*meta_temporal_suvr*Age',cross_cov,sep = "+")

# perform multiple linear regression:
lm(formula = mo3.3.3.4,data = data,contrasts = list(Gender = "contr.sum",Educ_Group=c(-1,1)))
```

- Examine whether there exists difference in brain system segregation between pathology groups:

```{r model 3.3.3.5}
# DV: whole-brain system segregation
# IV: Pathology Status (Aβ-/tau-, Aβ+/tau-,Aβ+/tau+)

model3.3.3.5 = paste('wb_seg ~ Patho_Group',cross_cov,sep = "+")  

# perform ANCOVA 
car::Anova(lm(formula = model3.3.3.5,data = data,contrasts=
              list(Patho_Group="contr.treatment",Gender = "contr.sum",Educ_Factor=c(-1,1))),type=3)
```

4) Examine the interaction between tau burden and APOE4 status on brain system segregation:

```{r model 3.3.3.6}
# DV: whole-brain system segregation
# IV: interaction between APOE4 status (APOE4+ vs. APOE4-) and meta-temporal tau SUVR

mod3.3.3.6 = paste('wb_seg ~ APOE4_group*meta_temporal_suvr',cross_cov,sep = "+")

lm(formula = mod3.3.3.6,data = data,contrasts = list(APOE4_group = "contr.sum",Gender = "contr.sum",Educ_Group=c(-1,1)))
```

### 3.3.4 Brain atrophy fully mediates the association between tau burden and brain system segregation

Examine whether neurodegeneration (cortical thickness & hippocampal volume) mediate the relationship between tau and brain system segregation:

```{r model 3.3.4.1}
# DV: whole-brain system segregation
# mediators: mean grey matter cortical thickness and mean hippocampal volume 

mod3.3.4.1 = 'wb_seg ~ meta_temporal_suvr-Age-Gender-Educ_Group-FD_postscrub+(hippo_vol)+(wb_mean_thickness)'

# perform mediation analysis
psych::mediate(mod3.3.4.1, data =stats_df,n.iter=1000)
```

### 3.3.5 Changes in AD-related pathology are not associated with changes in brain system segregation over time

1. Examine whether the change in amyloid is associated with change in brain system segregation over time: 

1) Examine the interaction between change in Centiloids and time on brain system segregation:
 
```{r model 3.3.5.1}
# DV: whole-brain system segregation
# IV: interaction between change in centiloids (i.e.,delta_centiloids) and time (normalized to baseline)

mod3.3.5.1 = paste('wb_seg ~ delta_centiloids*time',long_cov,"(1|Subject_ID)",sep = "+")

# perform linear-mixed effect regression
lmer(formula = model1.8,data = data,contrasts=list(Gender = "contr.sum",Educ_Group=c(-1,1)))
```

2) Examine the interaction between change in Centiloids, time and baseline on brain system segregation:

```{r model 3.3.5.2}
# DV: whole-brain system segregation
# IV: interaction between age at baseline, change in centiloids (i.e.,delta_centiloids) and time (normalized to baseline)
# the covariates for this model specifically is gender,education status, and head motion at baseline

mod3.3.5.2 = paste('wb_seg ~ age_bl*delta_centiloids*time',long_cov,"(1|Subject_ID)",sep = "+")

# perform linear-mixed effect regression
lmer(formula = mod3.3.5.2,data = data,contrasts=list(Gender = "contr.sum",Educ_Group=c(-1,1)))

```

2. Examine whether the change in tau burden is associated with change in brain system segregation over time: 

1) Examine the interaction between change in meta-temporal tau and time on brain system segregation:
 
```{r model 3.3.5.3}
# DV: whole-brain system segregation
# IV: interaction between change in meta-temporal tau burden (i.e.,delta_meta_temporal_suvr) and time (normalized to baseline)

mod3.3.5.3 = paste('wb_seg ~ delta_meta_temporal_suvr*time',long_cov,"(1|Subject_ID)",sep = "+")

# perform linear-mixed effect regression
lmer(formula = mod3.3.5.3,data = data,contrasts=list(Gender = "contr.sum",Educ_Group=c(-1,1)))
```

2) Examine the interaction between change in meta-temporal tau, time and baseline age on brain system segregation:

```{r model 3.3.5.4}
# DV: whole-brain system segregation
# IV: interaction between change in meta-temporal tau burden (i.e.,delta_meta_temporal_suvr), time (normalized to baseline) and age at baseline

mod3.3.5.4 = paste('wb_seg ~ delta_meta_temporal_suvr*time*Age_bl',long_cov,"(1|Subject_ID)",sep = "+")

# perform linear-mixed effect regression
lmer(formula = mod3.3.5.4,data = data,contrasts=list(Gender = "contr.sum",Educ_Group=c(-1,1)))

```


3) Examine whether the interaction between change in amyloid and change in tau burden is associated with change in brain system segregation:

```{r model 3.3.5.5}
# DV: whole-brain system segregation
# IV: interaction between change in meta-temporal tau burden (delta_meta_temporal_suvr), change in centiloids (delta_centiloids) and time (normalized to baseline)

mod3.3.5.5 = paste('wb_seg ~ delta_meta_temporal_suvr*delta_centiloids*time',long_cov,"(1|Subject_ID)",sep = "+")

# perform linear-mixed effect regression
lmer(formula = mod3.3.5.5,data = data,contrasts=list(Gender = "contr.sum",Educ_Group=c(-1,1)))

```




